### 2022-03-24 preprint.tex
Deploy to GitHub Pages

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

@@

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=4.5cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

@@

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\thispagestyle{empty}
{\Large\bfseries\sffamily Maximum entropy modelling of food-web
structure}
\vskip 5em

%
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,2,3}\quad %
\href{https://orcid.org/0000-0002-4498-7076}{Dominique\,Gravel}%
%
\,\textsuperscript{2,3}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,3}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Université de
Sherbrooke\quad \textsuperscript{3}\,Quebec Centre for Biodiversity
Science

\textbf{Correspondance to:}\\
Francis Banville --- \texttt{francis.banville@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
\textbf{\sffamily Abstract: }The principle of maximum entropy is a
rigorous mathematical method of finding constrained probability
distributions that has been proven useful in many ecological modelling
problems. However, despite its broad application in graph and network
theory, it has seldom been used to model ecological networks. Here we
show how the (joint) degree distribution of maximum entropy can be
directly derived using the number of species and the number of links in
terrestrial and aquatic food webs. We also present a heuristic and
flexible approach of finding the network of maximum entropy based on
simulating annealing and SVD entropy. We built two of these
network-level models using constraints given by the connectance and the
joint degree sequence, respectively. All maximum entropy models were
compared against open access food-web data and null and neutral models
commonly used in network ecology. We found that the maximum entropy
network model constrained by the joint degree sequence was a good
predictor of food-web structure, including nestedness and motifs
distribution. Overall, our results suggest that many properties of
ecological networks are mainly driven by the joint degree distribution
and statistical phenomena.
\vfill

\clearpage
\linenumbers
\pagestyle{normal}

>>>>

%!TEX TS-program = xelatex
\documentclass[10pt,oneside]{article}

@@

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}

\geometry{includemp,
            letterpaper,
            top=1.2in,
            bottom=2.510cm,
            inner=0.5in,
            outer=0.4in,
            marginparwidth=1.95in,
            marginparsep=0.4in}

\usepackage[singlelinecheck=off]{caption}
\captionsetup{
  font={small},
  labelfont={bf},
  format=plain,
  labelsep=quad
}
\usepackage{floatrow}
\floatsetup[figure]{margins=hangright,
              facing=no,
              capposition=beside,
              capbesideposition={center,outside},
              floatwidth=\textwidth}
\floatsetup[table]{margins=hangoutside,
             facing=yes,
             capposition=beside,
             capbesideposition={center,outside},
             floatwidth=\textwidth}

@@

\setcounter{secnumdepth}{5}

\usepackage{titlesec}

\titleformat{\section}[block]
{\normalfont\large\sffamily}
{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}

\titleformat{\subsection}[runin]
{\normalfont\fontseries{b}\selectfont\filright\sffamily}
{\thesubsection.}{.5em}{}

\titleformat{\subsubsection}[runin]
{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}

\fancypagestyle{firstpage}
{
   \fancyhf{}
   \renewcommand{\headrulewidth}{0pt}
   \fancyfoot[R]{\footnotesize\ccby}
   \fancyfoot[L]{\footnotesize\sffamily\today}
}

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}

\usepackage{tikz}
\begin{document}
\tikz [remember picture, overlay] %
\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
\pagestyle{normal}
\thispagestyle{firstpage}

\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}

\noindent {\LARGE \textbf{\textsf{Maximum entropy modelling of food-web
structure}}}

\medskip
\begin{flushleft}
{\small
%
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,2,3}, %
\href{https://orcid.org/0000-0002-4498-7076}{Dominique\,Gravel}%
%
\,\textsuperscript{2,3}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,3}
\vskip 1em
\textsuperscript{1}\,Université de
Montréal; \textsuperscript{2}\,Université de
Sherbrooke; \textsuperscript{3}\,Quebec Centre for Biodiversity
Science\\
\vskip 1em
\textbf{Correspondance to:}\\
Francis Banville --- \texttt{francis.banville@umontreal.ca}\\
}
\end{flushleft}

\vskip 2em
\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
\vskip 2em
\noindent

\marginpar{\vskip 1em\flushright
{\small{\bfseries Keywords}:\par
ecological modelling\\ecological networks\\food webs\\maximum
entropy\\null models\\}
}

\textbf{Abstract}:\,The principle of maximum entropy is a rigorous
mathematical method of finding constrained probability distributions
that has been proven useful in many ecological modelling problems.
However, despite its broad application in graph and network theory, it
has seldom been used to model ecological networks. Here we show how the
(joint) degree distribution of maximum entropy can be directly derived
using the number of species and the number of links in terrestrial and
aquatic food webs. We also present a heuristic and flexible approach of
finding the network of maximum entropy based on simulating annealing and
SVD entropy. We built two of these network-level models using
constraints given by the connectance and the joint degree sequence,
respectively. All maximum entropy models were compared against open
access food-web data and null and neutral models commonly used in
network ecology. We found that the maximum entropy network model
constrained by the joint degree sequence was a good predictor of
food-web structure, including nestedness and motifs distribution.
Overall, our results suggest that many properties of ecological networks
are mainly driven by the joint degree distribution and statistical
phenomena.

\vskip 2em
\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
\vskip 2em
